\textbf{DiracDec C++/Coq Implementation\cite{ExampleCitation}}

### 2024-12-08 main.tex
typing system updated. now using type index

\newcommand{\Type}{\mathsf{Type}}
\newcommand{\BaseS}{\mathsf{Base}}

@@

\section{TODO}
\begin{itemize}
    \item The Coq side should be able to decide the alpha-equivalence of terms.
\end{itemize}

\section{Language Syntax}
\begin{definition}[syntax]
    The syntax for Dirac notation types is defined as 
    \begin{align*}
        T ::=\ & x\ |\ \SType\ |\ \KType(\sigma)\ |\ \BType(\sigma)\ |\ \OType(\sigma_1, \sigma_2)\ |\ T_1 \to T_2\ |\ \SET(\sigma), \\
        \sigma ::=\ & x\ |\ \sigma_1 \times \sigma_2.
    \end{align*}
    The syntax for Dirac notation terms is defined as
    \begin{align*}
        e ::=\ & x\ |\ (e_1, e_2) \\
        & |\ 0\ |\ 1\ |\ \ADDS(e_1 \cdots e_n) \ |\ e_1 \times \cdots \times e_n\ |\ e^* \ |\ \delta_{e_1, e_2}\ |\ \DOT(e_1\ e_2) \\
        & |\ \ZEROK(\sigma)\ |\ \ZEROB(\sigma)\ |\ \ZEROO(\sigma_1, \sigma_2)\ |\ \ONEO(\sigma) \\
        & |\ \ket{e}\ |\ \bra{t}\ |\ e^\dagger\ |\ e_1.e_2\ |\ \ADD(e_1 \cdots e_n)\ |\ e_1 \otimes e_2 \\
        & |\ \MULK(e_1\ e_2)\ |\ \MULB(e_1\ e_2)\ |\ \OUTER(e_1\ e_2)\ |\ \MULO(e_1\ e_2) \\
        & |\ \mathbf{U}(e)\ |\ e_1 \star e_2\ |\ \sum_{e_1} e_2.
    \end{align*}
\end{definition}
 % after consideration I still think keep DOT as binary is the best choice. It reduces the unstability of complicated matching in sequences.

@@

\section{Typing System}
The type checking of Dirac notations involves maintaining a well-formed context $\Gamma$, which specifies the definitions and typing assumtpions for variables.
Notice that the type checking $\Gamma \vdash t : T$ is decidable.
\begin{figure}[h]
    \begin{gather*}
        \textbf{W-Empty} \qquad
        \frac{}{\WF([])}
        \qquad \qquad
        \textbf{W-Assum-Type} \qquad
        \frac{\WF(\Gamma) \qquad T \notin \Gamma}{\WF(\Gamma; T:\Type)}\\
        \\
        \textbf{W-Assum-Term} \qquad
        \frac{\Gamma\vdash T:\Type \qquad x \notin \Gamma}{\WF(\Gamma; x:T)}
        \qquad \qquad
        \textbf{W-Def} \qquad
        \frac{\Gamma\vdash t:T \qquad x \notin \Gamma}{\WF(\Gamma; x:=t:T)} \\
        \\
        \textbf{Var} \qquad
        \frac{\WF(\Gamma) \qquad (x:T) \in \Gamma \text{ or } (x := t : T) \in \Gamma \text{ for some $t$}}{\Gamma \vdash x : T} \\
        \\
        \textbf{Arrow} \qquad
        \frac{\Gamma \vdash T:\Type \qquad \Gamma \vdash U:\Type}{\Gamma \vdash T \to U : \Type} \\
        \\
        \textbf{Lam} \qquad
        \frac{\Gamma;x:T \vdash t : U}{\Gamma \vdash (\lambda x:T . t) : T \to U}
        \qquad \qquad
        \textbf{App} \qquad
        \frac{\Gamma \vdash t:U \to T \qquad \Gamma \vdash u:U}{\Gamma \vdash (t\ u):T}
    \end{gather*}
    \caption{The typing rules for the typed lambda calculus with environment.}
\end{figure}
\yx{The Lam rule is problematic. It still allows a ill-formed dependent type.}

\begin{figure}[h]
    \begin{gather*}
        \textbf{Type-Base} \qquad
        \frac{\WF(\Gamma)}{\Gamma \vdash \BaseS : \Type} 
        \qquad \qquad
        \textbf{Base-Subtype} \qquad
        \frac{\Gamma \vdash T : \BaseS}{\Gamma \vdash T : \Type} \\
        \\
        \textbf{Type-Ket} \qquad
        \frac{\Gamma \vdash \sigma : \BaseS}{\Gamma \vdash \KType(\sigma) : \Type}
        \qquad \qquad
        \textbf{Type-Bra} \qquad
        \frac{\Gamma \vdash \sigma : \BaseS}{\Gamma \vdash \BType(\sigma) : \Type} \\
        \\
        \textbf{Type-Opt} \qquad
        \frac{\Gamma \vdash \sigma : \BaseS \qquad \Gamma \vdash \tau : \BaseS}{\Gamma \vdash \OType(\sigma, \tau) : \Type}
        \qquad \qquad
        \textbf{Type-Scalar} \qquad
        \frac{\WF(\Gamma)}{\Gamma \vdash \SType : \Type} \\
        \\
        \textbf{Type-Set} \qquad
        \frac{\Gamma \vdash \sigma : \BaseS}{\Gamma \vdash \SET(\sigma) : \Type}
    \end{gather*}
    \caption{Typing rules for Dirac notation types.}
\end{figure}

\begin{figure}[h]
    \begin{gather*}
        \textbf{Sca-0} \qquad
        \frac{\WF(\Gamma)}{\Gamma \vdash 0 : \SType}
        \qquad \qquad
        \textbf{Sca-1} \qquad
        \frac{\WF(\Gamma)}{\Gamma \vdash 1 : \SType}
        \qquad \qquad
        \textbf{Sca-Delta} \qquad
        \frac{            
            \begin{aligned}
                & \Gamma\vdash \sigma : \BaseS \\
                & \Gamma\vdash s : \sigma \qquad \Gamma\vdash t : \sigma
            \end{aligned}} {\Gamma \vdash \delta_{s, t} : \SType} \\
        \\
        \textbf{Sca-Add} \qquad
        \frac{\Gamma\vdash a_i : \SType \text{ for all $i$}}{\Gamma\vdash a_1 + \cdots + a_n : \SType}
        \qquad \qquad
        \textbf{Sca-Mul} \qquad
        \frac{\Gamma\vdash a_i : \SType \text{ for all $i$}}{\Gamma\vdash a_1 \times \cdots \times a_n : \SType} \\
        \\
        \textbf{Sca-Conj} \qquad
        \frac{\Gamma \vdash a : \SType}{\Gamma \vdash a^*:\SType}
        \qquad \qquad
        \textbf{Sca-Dot} \qquad
        \frac{\Gamma\vdash B : \BType(\sigma) \qquad \Gamma\vdash K : \KType(\sigma)}{\Gamma \vdash B \cdot K : \SType}
    \end{gather*}
    \caption{Scalar typing rules.}
\end{figure}

\begin{figure}[h]
    \begin{gather*}
        % this two rules can be considered later
        % \textbf{Base-Def} \qquad
        % \frac{\WF(\Gamma) \qquad a_i \notin \Gamma \text{ for all $i$}}
        % {\Gamma\vdash \{a_1, a_2, \cdots, a_n\} : \BaseS}
        % \qquad \qquad
        % \textbf{Base-Element} \qquad
        % \frac{\WF(\Gamma) \qquad c \in \{a_1, a_2, \cdots a_n\}}{\Gamma\vdash c : \{a_1, a_2, \cdots, a_n\}} \\
        % \\
        \textbf{Base-Prod} \qquad
        \frac{\Gamma\vdash \sigma : \BaseS \qquad \Gamma\vdash \tau : \BaseS}{\Gamma\vdash \sigma \times \tau : \BaseS}
        \qquad \qquad
        \textbf{Pair-Base} \qquad
        \frac{
            \begin{aligned}
                & \Gamma\vdash \sigma : \BaseS \qquad \Gamma\vdash s : \sigma \\
                & \Gamma\vdash \tau : \BaseS \qquad \Gamma\vdash t : \tau
            \end{aligned}} {\Gamma\vdash (s, t) : \sigma \times \tau} 
        \textbf{}
    \end{gather*}
    \caption{Typing rules for bases. }
\end{figure}

\begin{figure}[h]
    \begin{gather*}
        \textbf{Ket-0} \qquad
        \frac{\Gamma \vdash \sigma : \BaseS}{\Gamma \vdash \ZEROK(\sigma) : \KType(\sigma)} 
        \qquad \qquad
        \textbf{Ket-Base} \qquad
        \frac{\Gamma \vdash \sigma : \BaseS \qquad \Gamma\vdash t : \sigma}{\Gamma \vdash \ket{t} : \KType(\sigma)} \\
        \\
        \textbf{Ket-Adj} \qquad
        \frac{\Gamma \vdash B : \BType(\sigma)}{\Gamma \vdash B^\dagger : \KType(\sigma)} 
        \qquad \qquad
        \textbf{Ket-Scr} \qquad
        \frac{\Gamma \vdash a : \SType \qquad \Gamma \vdash K : \KType(\sigma)}{\Gamma \vdash a.K : \KType(\sigma)} \\
        \\
        \textbf{Ket-Add} \qquad
        \frac{\Gamma \vdash K_i : \KType(\sigma) \text{ for all $i$}}{\Gamma \vdash K_1 + \cdots + K_n : \KType(\sigma)}
        \qquad \qquad
        \textbf{Ket-MulK} \qquad
        \frac{\Gamma \vdash O : \OType(\sigma, \tau) \qquad \Gamma \vdash K : \KType(\tau)}{\Gamma \vdash O \cdot K : \KType(\sigma)} \\
        \\
        \textbf{Ket-Tsr} \qquad
        \frac{\Gamma \vdash K_1 : \KType(\sigma) \qquad \Gamma \vdash K_2 : \KType(\tau)} {\Gamma \vdash K_1 \otimes K_2 : \KType(\sigma \times \tau)}
    \end{gather*}
    \caption{Ket typing rules.}
\end{figure}

\begin{figure}[h]
    \begin{gather*}
        \textbf{Bra-0} \qquad
        \frac{\Gamma \vdash \sigma : \BaseS}{\Gamma \vdash \ZEROB(\sigma) : \BType(\sigma)} 
        \qquad \qquad
        \textbf{Bra-Base} \qquad
        \frac{\Gamma \vdash \sigma : \BaseS \qquad \Gamma\vdash t : \sigma}{\Gamma \vdash \bra{t} : \BType(\sigma)} \\
        \\
        \textbf{Bra-Adj} \qquad
        \frac{\Gamma \vdash K : \KType(\sigma)}{\Gamma \vdash K^\dagger : \BType(\sigma)} 
        \qquad \qquad
        \textbf{Bra-Scr} \qquad
        \frac{\Gamma \vdash a : \SType \qquad \Gamma \vdash B : \BType(\sigma)}{\Gamma \vdash a.B : \BType(\sigma)} \\
        \\
        \textbf{Bra-Add} \qquad
        \frac{\Gamma \vdash B_i : \BType(\sigma) \text{ for all $i$}}{\Gamma \vdash B_1 + \cdots + B_n : \BType(\sigma)}
        \qquad \qquad
        \textbf{Bra-MulB} \qquad
        \frac{\Gamma \vdash B : \KType(\sigma) \qquad \Gamma \vdash O : \OType(\sigma, \tau)}{\Gamma \vdash B \cdot O : \BType(\tau)} \\
        \\
        \textbf{Bra-Tsr} \qquad
        \frac{\Gamma \vdash B_1 : \BType(\sigma) \qquad \Gamma \vdash B_2 : \BType(\tau)} {\Gamma \vdash B_1 \otimes B_2 : \BType(\sigma \times \tau)}
    \end{gather*}
    \caption{Bra typing rules.}
\end{figure}

\begin{figure}[h]
    \begin{gather*}
        \textbf{Opt-0} \qquad
        \frac{\Gamma \vdash \sigma : \BaseS \qquad \Gamma \vdash \tau : \BaseS}{\Gamma \vdash \ZEROO(\sigma, \tau) : \OType(\sigma, \tau)} 
        \qquad \qquad
        \textbf{Opt-1} \qquad
        \frac{\Gamma \vdash \sigma : \BaseS}{\Gamma \vdash \ONEO(\sigma) : \OType(\sigma, \sigma)} \\
        \\
        \textbf{Opt-Adj} \qquad
        \frac{\Gamma \vdash O : \OType(\sigma, \tau)}{\Gamma \vdash O^\dagger : \OType(\tau, \sigma)} 
        \qquad \qquad
        \textbf{Opt-Scr} \qquad
        \frac{\Gamma \vdash a : \SType \qquad \Gamma \vdash O : \OType(\sigma, \tau)}{\Gamma \vdash a.O : \OType(\sigma, \tau)} \\
        \\
        \textbf{Opt-Add} \qquad
        \frac{\Gamma \vdash O_i : \OType(\sigma, \tau) \text{ for all $i$}}{\Gamma \vdash O_1 + \cdots + O_n : \OType(\sigma, \tau)}
        \qquad \qquad
        \textbf{Opt-Outer} \qquad
        \frac{\Gamma\vdash K : \KType(\sigma) \qquad \Gamma\vdash B : \BType(\tau)}{\Gamma\vdash K \cdot B : \OType(\sigma, \tau)} \\
        \\
        \textbf{Opt-MulO} \qquad
        \frac{\Gamma \vdash O_1 : \OType(\sigma, \tau) \qquad \Gamma \vdash O_2 : \OType(\tau, \rho)}{\Gamma \vdash O_1 \cdot O_2 : \OType(\sigma, \rho)} \\
        \\
        \textbf{Opt-Tsr} \qquad
        \frac{\Gamma \vdash O_1 : \OType(\sigma_1, \tau_1) \qquad \Gamma \vdash O_2 : \OType(\sigma_2, \tau_2)} {\Gamma \vdash O_1 \otimes O_2 : \OType(\sigma_1 \times \sigma_2, \tau_1 \times \tau_2)}
    \end{gather*}
    \caption{Operator typing rules.}
\end{figure}

\begin{figure}[h]
    \begin{gather*}
        \textbf{Set-U} \qquad
        \frac{\Gamma \vdash \sigma : \BaseS}{\Gamma \vdash \mathbf{U}(\sigma) : \SET(\sigma)} 
        \qquad \qquad
        \textbf{Set-Prod} \qquad
        \frac{\Gamma \vdash A : \SET(\sigma) \qquad \Gamma \vdash B : \SET(\tau)}{\Gamma \vdash A \star B : \SET(\sigma \times \tau)}
    \end{gather*}
    \caption{Set typing rules.}
\end{figure}

\begin{figure}[h]
    \begin{gather*}
        \textbf{Sum-Scalar} \qquad
        \frac{\Gamma \vdash s : \SET(\sigma) \qquad \Gamma \vdash f : \sigma \to \SType}{\Gamma \vdash \sum_{s} f : \SType} 
        \qquad \qquad
        \textbf{Sum-Ket} \qquad
        \frac{\Gamma \vdash s : \SET(\sigma) \qquad \Gamma \vdash f : \sigma \to \KType(\tau)}{\Gamma \vdash \sum_{s} f : \KType(\tau)} \\
        \\
        \textbf{Sum-Bra} \qquad
        \frac{\Gamma \vdash s : \SET(\sigma) \qquad \Gamma \vdash f : \sigma \to \BType(\tau)}{\Gamma \vdash \sum_{s} f : \BType(\tau)} 
        \qquad \qquad
        \textbf{Sum-Opt} \qquad
        \frac{\Gamma \vdash s : \SET(\sigma) \qquad \Gamma \vdash f : \sigma \to \OType(\tau, \rho)}{\Gamma \vdash \sum_{s} f : \OType(\tau, \rho)}
    \end{gather*}
    \caption{Sum typing rules.}
\end{figure}

@@

\yx{Alpha rule will be implemented by renaming, and we can leave it for the future.}

>>>>

\newcommand{\Index}{\mathsf{Index}}
\newcommand{\Type}{\mathsf{Type}}
\newcommand{\Basis}{\mathsf{Basis}}

@@

\section{TODO}
\begin{itemize}
    \item Alpha equivalence is checked by a bound variable renaming on the whole term.
    \item arrow type of $\Type \to \Type$ are forbidden. Check whether typing rules comply with this restriction.
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Language Syntax}
\begin{definition}[syntax]
    The syntax for type indices are defined as
    \begin{align*}
        \sigma ::=\ & x\ |\ \sigma_1 \times \sigma_2.
    \end{align*}
    The syntax for Dirac notation types is defined as 
    \begin{align*}
        T ::=\ & x\ |\ \Basis(\sigma)\ |\ \SType\ |\ \KType(\sigma)\ |\ \BType(\sigma)\ |\ \OType(\sigma_1, \sigma_2)\ |\ T_1 \to T_2\ |\ \forall x.T\ |\ \SET(\sigma). \\
    \end{align*}
    The syntax for Dirac notation terms is defined as
    \begin{align*}
        e ::=\ & x\ |\ (e_1, e_2) \\
        & |\ 0\ |\ 1\ |\ \ADDS(e_1 \cdots e_n) \ |\ e_1 \times \cdots \times e_n\ |\ e^* \ |\ \delta_{e_1, e_2}\ |\ \DOT(e_1\ e_2) \\
        & |\ \ZEROK(\sigma)\ |\ \ZEROB(\sigma)\ |\ \ZEROO(\sigma_1, \sigma_2)\ |\ \ONEO(\sigma) \\
        & |\ \ket{e}\ |\ \bra{t}\ |\ e^\dagger\ |\ e_1.e_2\ |\ \ADD(e_1 \cdots e_n)\ |\ e_1 \otimes e_2 \\
        & |\ \MULK(e_1\ e_2)\ |\ \MULB(e_1\ e_2)\ |\ \OUTER(e_1\ e_2)\ |\ \MULO(e_1\ e_2) \\
        & |\ \mathbf{U}(e)\ |\ e_1 \star e_2\ |\ \sum_{e_1} e_2.
    \end{align*}
\end{definition}
 % after consideration I still think keep DOT as binary is the best choice. It reduces the unstability of complicated matching in sequences.

@@

\section{Typing System}
The type checking of Dirac notations involves maintaining a well-formed context $\Gamma$, which specifies the definitions and typing assumtpions for variables.
Notice that the type checking $\Gamma \vdash t : T$ is decidable.

\begin{definition}[context]
    \begin{align*}
        \Gamma ::= []\ |\ \Gamma; x : \Index\ |\ \Gamma; T : \Type\ |\ \Gamma; x : T\ |\ \Gamma; x := t : T.
    \end{align*}
\end{definition}

\begin{figure}[h]
    \begin{gather*}
        \textbf{W-Empty} \qquad
        \frac{}{\WF([])} \\
        \\
        \textbf{W-Assum-Index} \qquad
        \frac{\WF(\Gamma) \qquad \sigma \notin \Gamma}{\WF(\Gamma; \sigma : \Index)}\qquad \qquad
        \textbf{W-Assum-Type} \qquad
        \frac{\WF(\Gamma) \qquad T \notin \Gamma}{\WF(\Gamma; T : \Type)} \\
        \\
        \textbf{W-Assum-Term} \qquad
        \frac{\Gamma \vdash T : \Type \qquad x \notin \Gamma}{\WF(\Gamma; x:T)}
        \qquad \qquad
        \textbf{W-Def-Term} \qquad
        \frac{\Gamma\vdash t:T \qquad x \notin \Gamma}{\WF(\Gamma; x:=t:T)}
    \end{gather*}
    \caption{Rules for a well-formed context.}
\end{figure}

\begin{figure}
    \begin{gather*}
        \textbf{Index-Var} \qquad
        \frac{\WF(\Gamma) \qquad \sigma : \Index \in \Gamma}{\Gamma \vdash \sigma : \Index}
        \qquad \qquad
        \textbf{Index-Prod} \qquad
        \frac{\Gamma \vdash \sigma : \Index \qquad \Gamma \vdash \tau : \Index}{\Gamma \vdash \sigma \times \tau : \Index}
    \end{gather*}
    \caption{Rules for type index.}
\end{figure}

\begin{figure}
    \begin{gather*}
        \textbf{Type-Var} \qquad
        \frac{\WF(\Gamma) \qquad T : \Type \in \Gamma}{\Gamma \vdash T : \Type}\\
        \\
        \textbf{Type-Arrow} \qquad
        \frac{\Gamma \vdash T : \Type \qquad \Gamma \vdash U : \Type}{\Gamma \vdash T \to U : \Type}
        \qquad \qquad
        \textbf{Type-Index} \qquad
        \frac{\Gamma ; \sigma : \Index \vdash U : \Type}{\Gamma \vdash \forall \sigma.U : \Type} \\
        \\        
        \textbf{Type-Basis} \qquad
        \frac{\Gamma \vdash \sigma : \Index}{\Gamma \vdash \Basis(\sigma) : \Type}  \\
        \\
        \textbf{Type-Ket} \qquad
        \frac{\Gamma \vdash \sigma : \Index}{\Gamma \vdash \KType(\sigma) : \Type}
        \qquad \qquad
        \textbf{Type-Bra} \qquad
        \frac{\Gamma \vdash \sigma : \Index}{\Gamma \vdash \BType(\sigma) : \Type} \\
        \\
        \textbf{Type-Opt} \qquad
        \frac{\Gamma \vdash \sigma : \Index \qquad \Gamma \vdash \tau : \Index}{\Gamma \vdash \OType(\sigma, \tau) : \Type}
        \qquad \qquad
        \textbf{Type-Scalar} \qquad
        \frac{\WF(\Gamma)}{\Gamma \vdash \SType : \Type} \\
        \\
        \textbf{Type-Set} \qquad
        \frac{\Gamma \vdash \sigma : \Index}{\Gamma \vdash \SET(\sigma) : \Type}
    \end{gather*}
    \caption{Rules for types.}
\end{figure}

\begin{figure}[h]
    \begin{gather*}
        \textbf{Term-Var} \qquad
        \frac{\WF(\Gamma) \qquad (x:T) \in \Gamma \text{ or } (x := t : T) \in \Gamma \text{ for some $t$}}{\Gamma \vdash x : T} \\
        \\
        \textbf{Lam} \qquad
        \frac{\Gamma;x:T \vdash t : U}{\Gamma \vdash (\lambda x:T . t) : T \to U}
        \qquad \qquad
        \textbf{Index} \qquad
        \frac{\Gamma; \sigma : \Index \vdash t : U \qquad x \notin \Gamma}{\Gamma \vdash (\lambda x.t) : \forall \sigma.U} \\
        \\
        \textbf{App-Arrow} \qquad
        \frac{\Gamma \vdash t:U \to T \qquad \Gamma \vdash u:U}{\Gamma \vdash (t\ u):T}
        \qquad \qquad
        \textbf{App-Index} \qquad
        \frac{\Gamma \vdash t:\forall x.U \qquad \Gamma \vdash u : \Index}{\Gamma \vdash (t\ u):U\{x/u\}}
    \end{gather*}
    \caption{Rules for variable and function typings.}
\end{figure}

\begin{figure}[h]
    \begin{gather*}
        \textbf{Sca-0} \qquad
        \frac{\WF(\Gamma)}{\Gamma \vdash 0 : \SType}
        \qquad \qquad
        \textbf{Sca-1} \qquad
        \frac{\WF(\Gamma)}{\Gamma \vdash 1 : \SType}
        \qquad \qquad
        \textbf{Sca-Delta} \qquad
        \frac{ \Gamma\vdash s : \Basis(\sigma) \qquad \Gamma\vdash t : \Basis(\sigma) } {\Gamma \vdash \delta_{s, t} : \SType} \\
        \\
        \textbf{Sca-Add} \qquad
        \frac{\Gamma\vdash a_i : \SType \text{ for all $i$}}{\Gamma\vdash a_1 + \cdots + a_n : \SType}
        \qquad \qquad
        \textbf{Sca-Mul} \qquad
        \frac{\Gamma\vdash a_i : \SType \text{ for all $i$}}{\Gamma\vdash a_1 \times \cdots \times a_n : \SType} \\
        \\
        \textbf{Sca-Conj} \qquad
        \frac{\Gamma \vdash a : \SType}{\Gamma \vdash a^*:\SType}
        \qquad \qquad
        \textbf{Sca-Dot} \qquad
        \frac{\Gamma\vdash B : \BType(\sigma) \qquad \Gamma\vdash K : \KType(\sigma)}{\Gamma \vdash B \cdot K : \SType}
    \end{gather*}
    \caption{Scalar typing rules.}
\end{figure}

\begin{figure}[h]
    \begin{gather*}
        % this two rules can be considered later
        % \textbf{Base-Def} \qquad
        % \frac{\WF(\Gamma) \qquad a_i \notin \Gamma \text{ for all $i$}}
        % {\Gamma\vdash \{a_1, a_2, \cdots, a_n\} : \Basis}
        % \qquad \qquad
        % \textbf{Base-Element} \qquad
        % \frac{\WF(\Gamma) \qquad c \in \{a_1, a_2, \cdots a_n\}}{\Gamma\vdash c : \{a_1, a_2, \cdots, a_n\}} \\
        % \\
        \textbf{Pair-Base} \qquad
        \frac{
            \Gamma\vdash s : \Basis(\sigma) \qquad \Gamma\vdash t : \Basis(\tau)} {\Gamma\vdash (s, t) : \Basis(\sigma \times \tau)} 
        \textbf{}
    \end{gather*}
    \caption{Typing rules for Basis. }
\end{figure}

\begin{figure}[h]
    \begin{gather*}
        \textbf{Ket-0} \qquad
        \frac{\Gamma \vdash \sigma : \Index}{\Gamma \vdash \ZEROK(\sigma) : \KType(\sigma)} 
        \qquad \qquad
        \textbf{Ket-Base} \qquad
        \frac{\Gamma\vdash t : \Basis(\sigma)}{\Gamma \vdash \ket{t} : \KType(\sigma)} \\
        \\
        \textbf{Ket-Adj} \qquad
        \frac{\Gamma \vdash B : \BType(\sigma)}{\Gamma \vdash B^\dagger : \KType(\sigma)} 
        \qquad \qquad
        \textbf{Ket-Scr} \qquad
        \frac{\Gamma \vdash a : \SType \qquad \Gamma \vdash K : \KType(\sigma)}{\Gamma \vdash a.K : \KType(\sigma)} \\
        \\
        \textbf{Ket-Add} \qquad
        \frac{\Gamma \vdash K_i : \KType(\sigma) \text{ for all $i$}}{\Gamma \vdash K_1 + \cdots + K_n : \KType(\sigma)}
        \qquad \qquad
        \textbf{Ket-MulK} \qquad
        \frac{\Gamma \vdash O : \OType(\sigma, \tau) \qquad \Gamma \vdash K : \KType(\tau)}{\Gamma \vdash O \cdot K : \KType(\sigma)} \\
        \\
        \textbf{Ket-Tsr} \qquad
        \frac{\Gamma \vdash K_1 : \KType(\sigma) \qquad \Gamma \vdash K_2 : \KType(\tau)} {\Gamma \vdash K_1 \otimes K_2 : \KType(\sigma \times \tau)}
    \end{gather*}
    \caption{Ket typing rules.}
\end{figure}

\begin{figure}[h]
    \begin{gather*}
        \textbf{Bra-0} \qquad
        \frac{\Gamma \vdash \sigma : \Index}{\Gamma \vdash \ZEROB(\sigma) : \BType(\sigma)} 
        \qquad \qquad
        \textbf{Bra-Base} \qquad
        \frac{\Gamma\vdash t : \Basis(\sigma)}{\Gamma \vdash \bra{t} : \BType(\sigma)} \\
        \\
        \textbf{Bra-Adj} \qquad
        \frac{\Gamma \vdash K : \KType(\sigma)}{\Gamma \vdash K^\dagger : \BType(\sigma)} 
        \qquad \qquad
        \textbf{Bra-Scr} \qquad
        \frac{\Gamma \vdash a : \SType \qquad \Gamma \vdash B : \BType(\sigma)}{\Gamma \vdash a.B : \BType(\sigma)} \\
        \\
        \textbf{Bra-Add} \qquad
        \frac{\Gamma \vdash B_i : \BType(\sigma) \text{ for all $i$}}{\Gamma \vdash B_1 + \cdots + B_n : \BType(\sigma)}
        \qquad \qquad
        \textbf{Bra-MulB} \qquad
        \frac{\Gamma \vdash B : \KType(\sigma) \qquad \Gamma \vdash O : \OType(\sigma, \tau)}{\Gamma \vdash B \cdot O : \BType(\tau)} \\
        \\
        \textbf{Bra-Tsr} \qquad
        \frac{\Gamma \vdash B_1 : \BType(\sigma) \qquad \Gamma \vdash B_2 : \BType(\tau)} {\Gamma \vdash B_1 \otimes B_2 : \BType(\sigma \times \tau)}
    \end{gather*}
    \caption{Bra typing rules.}
\end{figure}

\begin{figure}[h]
    \begin{gather*}
        \textbf{Opt-0} \qquad
        \frac{\Gamma \vdash \sigma : \Index \qquad \Gamma \vdash \tau : \Index}{\Gamma \vdash \ZEROO(\sigma, \tau) : \OType(\sigma, \tau)} 
        \qquad \qquad
        \textbf{Opt-1} \qquad
        \frac{\Gamma \vdash \sigma : \Index}{\Gamma \vdash \ONEO(\sigma) : \OType(\sigma, \sigma)} \\
        \\
        \textbf{Opt-Adj} \qquad
        \frac{\Gamma \vdash O : \OType(\sigma, \tau)}{\Gamma \vdash O^\dagger : \OType(\tau, \sigma)} 
        \qquad \qquad
        \textbf{Opt-Scr} \qquad
        \frac{\Gamma \vdash a : \SType \qquad \Gamma \vdash O : \OType(\sigma, \tau)}{\Gamma \vdash a.O : \OType(\sigma, \tau)} \\
        \\
        \textbf{Opt-Add} \qquad
        \frac{\Gamma \vdash O_i : \OType(\sigma, \tau) \text{ for all $i$}}{\Gamma \vdash O_1 + \cdots + O_n : \OType(\sigma, \tau)}
        \qquad \qquad
        \textbf{Opt-Outer} \qquad
        \frac{\Gamma\vdash K : \KType(\sigma) \qquad \Gamma\vdash B : \BType(\tau)}{\Gamma\vdash K \cdot B : \OType(\sigma, \tau)} \\
        \\
        \textbf{Opt-MulO} \qquad
        \frac{\Gamma \vdash O_1 : \OType(\sigma, \tau) \qquad \Gamma \vdash O_2 : \OType(\tau, \rho)}{\Gamma \vdash O_1 \cdot O_2 : \OType(\sigma, \rho)} \\
        \\
        \textbf{Opt-Tsr} \qquad
        \frac{\Gamma \vdash O_1 : \OType(\sigma_1, \tau_1) \qquad \Gamma \vdash O_2 : \OType(\sigma_2, \tau_2)} {\Gamma \vdash O_1 \otimes O_2 : \OType(\sigma_1 \times \sigma_2, \tau_1 \times \tau_2)}
    \end{gather*}
    \caption{Operator typing rules.}
\end{figure}

\begin{figure}[h]
    \begin{gather*}
        \textbf{Set-U} \qquad
        \frac{\Gamma \vdash \sigma : \Index}{\Gamma \vdash \mathbf{U}(\sigma) : \SET(\sigma)} 
        \qquad \qquad
        \textbf{Set-Prod} \qquad
        \frac{\Gamma \vdash A : \SET(\sigma) \qquad \Gamma \vdash B : \SET(\tau)}{\Gamma \vdash A \star B : \SET(\sigma \times \tau)}
    \end{gather*}
    \caption{Set typing rules.}
\end{figure}

\begin{figure}[h]
    \begin{gather*}
        \textbf{Sum-Scalar} \qquad
        \frac{\Gamma \vdash s : \SET(\sigma) \qquad \Gamma \vdash f : \Basis(\sigma) \to \SType}{\Gamma \vdash \sum_{s} f : \SType} \\
        \\
        \textbf{Sum-Ket} \qquad
        \frac{\Gamma \vdash s : \SET(\sigma) \qquad \Gamma \vdash f : \Basis(\sigma) \to \KType(\tau)}{\Gamma \vdash \sum_{s} f : \KType(\tau)} \\
        \\
        \textbf{Sum-Bra} \qquad
        \frac{\Gamma \vdash s : \SET(\sigma) \qquad \Gamma \vdash f : \Basis(\sigma) \to \BType(\tau)}{\Gamma \vdash \sum_{s} f : \BType(\tau)} \\
        \\
        \textbf{Sum-Opt} \qquad
        \frac{\Gamma \vdash s : \SET(\sigma) \qquad \Gamma \vdash f : \Basis(\sigma) \to \OType(\tau, \rho)}{\Gamma \vdash \sum_{s} f : \OType(\tau, \rho)}
    \end{gather*}
    \caption{Sum typing rules.}
\end{figure}

@@
